---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
```

The Dataset
YERockfish.csv contains measurements of fish collected from along the Oregon coast

Yelloweye Rockfish (Sebastes rubberimus)
Length (length; to the nearest cm),

Age (years)

Maturity (Immature and Mature) state of female

Read in the data as a tibble
```{r}
fish_data = as_tibble(YERockfish, sep = "\t", header = TRUE)
fish_data
```

This data contains a date column. R provides an easy way to manipulate dates.

The field needs to be converted to a format that R will recognize as a date.
The as.POSIXct() function does this conversion. The function takes two arguments
The column to convert
Format of the date. Since dates come in vairous shapes (ex. yyyy/m/d, mm/dd/yy, etc..)
Use the function as.POSIXct() to convert the date. The approach is exactly the same as converting to factor (as.factor()) or integer (as.integer())

Here you can pass the following format string for the parameter format

%m/%d/%Y
Bonus: use a dplyr pipeline to mutate the column

```{r}
?add_column
```


```{r}
new_date <- as.Date(fish_data$date, "%m/%d/%Y")
new_fish_data <- add_column(fish_data, new_date, .before = 2) 
                  
fish_data_date = subset(new_fish_data, select = -c(date))

head(fish_data_date)
```


```{r}
correct_date = as.POSIXct(fish_data_date$new_date, "%Y-%m-%d")
correct_date

correct_date_data <- add_column(fish_data_date, correct_date, .before = 2)
correct_date_data

correct_fish_data = subset(correct_date_data, select = -c(new_date))
head(correct_fish_data)
```

Use the function head() on your modified tibble to make sure the column is now of type dttm (date and time) instead of char

```{r}
head(correct_fish_data)
```

Count the number of lines in your file

```{r}
count(correct_fish_data)
```

Plot the count of observations per year.

Note that you can easily parse the year from the data using the year() function, which takes a column of dttm datatype
The function year is part of the lubridate package, which you may need to import
Hint 1: try the function year() on your date column

Hint 2: Use group by to group the data and n() to count the number of entries in each group

Bonus: use a dplyr pipeline to answer this question.

```{r}
#Try year() on the date colum
date_year = year(correct_fish_data$correct_date)
date_year

#mutate, new column and filter by column
```
```{r}
#This method works
large_date_sample_data <- correct_fish_data %>%
  mutate(sample_year = year(correct_date)) %>%
  filter(sample_year != 2004, sample_year != 2008 )
large_date_sample_data
```
```{r}
?filter
```



```{r}
year_count <- correct_fish_data %>% 
    group_by(year(correct_fish_data$correct_date)) %>% 
    count()
year_count
```
```{r}
plot(year_count)
```



Remove all entries that below to a year for which there are less than 5 entries
E.g. there is only one entry for 2008, so we can remove it.
Save the data to a new tibble

```{r}
large_date_sample_data
```

```{r}
# WHY DOESN"T THIS WORK?!
correct_fish_data %>% 
  group_by(year(correct_fish_data$correct_date)) %>% 
  count() %>% 
  filter(n > 5) %>%
ungroup(year(correct_fish_data$correct_date))
```

```{r}
correct_fish_data %>% filter(correct_fish_data$correct_date, )
```


Count the number of entries and make sure there are less observations

```{r}
count(large_date_sample_data)
#made sure there were 4 less. Initial test showed that the pipe was only removing 2004 or 2008
```

Model the fish maturity using the fish length

i.e. predict maturity from the lengh data
You can make any changes to the data needed to build this model

```{r}
numeric_maturity = function(x) {
    if(x == "Mature"){
      1
    }else{
      0
    }
}
numeric_maturity("Mature")
numeric_maturity("Immature")
numeric_maturity("NA")
```

```{r}
maturity_tibble = tibble(large_date_sample_data$length, large_date_sample_data$age, large_date_sample_data$maturity)
maturity_tibble
```

```{r}
no_na_maturity = na.omit(maturity_tibble)
no_na_maturity
```


```{r}
numeric = mapply(numeric_maturity, no_na_maturity$`large_date_sample_data$maturity`)
numeric
```
```{r}
numeric_no_na = tibble(no_na_maturity, numeric)
numeric_no_na
```

```{r}
lm(numeric_no_na$numeric~numeric_no_na$`large_date_sample_data$length`)
```


```{r}
ggplot()+
  geom_point(aes(x = numeric_no_na$`large_date_sample_data$length`, y = numeric_no_na$numeric)) +
  geom_hline(yintercept = 0.5, linetype= "dashed", alpha=0.6, size = 1) +
  geom_abline(intercept = -0.52343, slope = 0.02874, color = "red", size = 1) +
  xlab("Length") +
  ylab ("Status") +
  theme(text = element_text(size = 15))
```


Generate a plot to show the data and the fit of the model (sigmoid)

```{r}
logistic_reg_model = glm(data = numeric_no_na, numeric~numeric_no_na$`large_date_sample_data$length`, family = "binomial")
summary(logistic_reg_model)
```

```{r}
logistic_reg_model$coefficients
```

```{r}
beta_0 = logistic_reg_model$coefficients[1]
beta_1 = logistic_reg_model$coefficients[2]

x_axis = seq(min(numeric_no_na$`large_date_sample_data$length`)-3, max(numeric_no_na$`large_date_sample_data$length`)+3, 0.05)
g_x = 1/(1+exp(-(beta_0 + beta_1 * x_axis)))

ggplot()+
  geom_point(aes(x = numeric_no_na$`large_date_sample_data$length`, y = numeric_no_na$numeric)) +
  geom_line(aes(x_axis, g_x)) +
  xlab("Length") +
  ylab("Maturity") +
  theme(text = element_text(size = 22))
```


What is the length at which the probability of picking a mature fish is 0.5?
You can eyeball it or can you compute it formally from the logistic regression

```{r}
# At about 38 the probability of picking a mature fish is 0.5. 
```

Add an era column to your dataset such that

era has the value "pre_2000" if the year of the observation is pre 2002
era has the value "era2002 and after" otherwise
Hint: Check the if_else in dplyr

https://dplyr.tidyverse.org/reference/if_else.html
Bonus: implement this operation using a pipeline and mutate

```{r}
era_2002 <- function(x){
    if(x < 2002){
      "pre_2000"
    }else{
      "era2002"
    }
}
era_2002("2000")
era_2002("1999")
era_2002("2002")
era_2002("2003")
```

```{r}
era <- mapply(era_2002, large_date_sample_data$sample_year)

```

```{r}
era_large_data <- large_date_sample_data %>%
  mutate(era, .keep = "all" )

era_large_data
```

You suspected that some major environmental stressor that occurred in 2002 had an impact on the fish length and maturity

Build a logistic regression for maturity as an outcome using era and length as predictive variables

Make sure your formula accounts for interactions
Use an ANOVA to test whether maturity is a function of both length and era

Does the maturity differ between the two eras? i.e. is the era model coefficient significant?

```{r}
era_maturity = era_large_data %>%
    mutate(-(correct_date))
```

Final note:
This practical is inspired from: http://derekogle.com/IFAR/supplements/maturity/index.html#fitting-the-logistic-regression-model

The link above contains most answers, so please do not read until you are done with this practical.
